Indirect elicitation as a necessary condition for consistent surrogate losses

### 2020-06-04 neurips-20/cvx-flats.tex
pass through Sec 3

We start by formalizing consistency in two ways that generalize across our four quadrants.
First, given an target loss $\ell$, then we say $L$ is consistent if optimizing $L$ implies optimizing $\ell$ (Definition~\ref{def:consistent-ell}).
Second, given a target property $\gamma$, such as the $\alpha$-quantile, we say $L$ is consistent if optimizing $L$ implies approaching, in some sense, the correct statistic $\gamma(D_x)$ of the conditional distributions $D_x = \Pr[Y|X=x]$ (Definition~\ref{def:consistent-prop}).
We then observe that Definition~\ref{def:consistent-ell} is subsumed by Definition~\ref{def:consistent-prop}, and use this to show consistency always implies $L$ indirectly elicits $\prop{\ell}$ or $\gamma$ respectively.

\begin{definition}[Consistent: loss]\label{def:consistent-ell}
	A loss and link $(L,\psi)$ are consistent with respect to an target loss $\ell$ if, for all distributions $D$ over input and label spaces $\X \times\Y$, and for all sequences of measurable hypothesis functions $\{f_m : \X \to \R\}$,
	\begin{align*}
	\E_D L(f_m(X), Y) \to \inf_f \E_D L(f(X), Y) &\implies \E_D \ell((\psi \circ f_m)(X), Y) \to \inf_f \E_D \ell((\psi \circ f)(X), Y)~.~
	\end{align*}
\end{definition}
%% Bo: resolved, Jun 1
%\botodo{On the right side, is measurable enough? I guess we need $\psi \circ f$ to be measurable...on the other hand, we don't want to constrain $f$ based on $\psi$. Do we need a condition on $\psi$ that guarantees $\psi \circ f$ is measurable?}
%\raft{I think okay as written -- if $\psi$ is restrictive, you won't achieve consistency.  And typically $f$ is much more expressive anyway, e.g. $\reals^\Y$ for classification.}

Instead of a target loss $\ell$, one may want to learn a conditional statistic such as the expected value, variance, or entropy.
In this case, following the tradition in the statistics literature on conditional estimation~\citep{gyorfi2006distribution,fan1998efficient,ruppert1997local},
we formalize consistency as converging to the correct conditional estimates of the property.
Convergence is measured by functions $\propdis(r, p)$ that formalize how close $r$ is to ``correct'' for conditional distribution $p$.
In particular we should have $\propdis(r,p) = 0 \iff r \in \gamma(p)$.
\btw{Bo: Would be nice to give some natural special cases: for a finite property, $\propdis(r,p) = \mathbf{1}[r \in \gamma(p)]$, and for single-valued properties with a distance metric on $\R$, $\propdis(r,p) = \text{dist}(r, \gamma(p))$.}

@@

When noting that some properties are not elicitable, including the variance, Lemma~\ref{lem:consistent-loss-implies-prop} confirms that consistency with respect to a property is a strictly broader notion, and further points to indirect elicitation as a natural necessary condition for consistency.
We now show this necessity directly.

\begin{theorem}\label{thm:consistent-implies-indir-elic}
	For a surrogate $L \in \L$, if the pair $(L, \psi)$ is consistent with respect to a property $\gamma$ or loss $\ell$ eliciting $\gamma$, then $(L, \psi)$ indirectly elicits $\gamma$.
\end{theorem}
\begin{proof}
%  \raft{Maybe ref the consistency definition here explicitly, at the top and the end}
	By Lemma~\ref{lem:consistent-loss-implies-prop}, it suffices to show the result for consistency with respect to a property $\gamma$, setting $\gamma := \prop{\ell}$ if $\ell$ is given instead.
	We show the contrapositive; suppose $(L, \psi)$ does not indirectly elicit $\gamma$, meaning we have some $p \in \simplex$ so that $u \in \Gamma(p)$ but $\psi(u) \not \in \gamma(p)$, where $\Gamma := \prop{L}$.
	Observe that we use the fact $\Gamma(p) \neq \emptyset$.
	Consider the constant sequence $\{f_m\}$ with $f_m(x) = u$ for all $m,x$, and take $D$ with full support on some $x \in \X$, and let $D_x = p$.
  Since $u \in \Gamma(p)$, we observe $\E_D L(f_m(X), Y) = \inf_f L(f(X),Y)$ for all $m$; in particular $\E_D L(f_m(X), Y) \to \inf_f L(f(X),Y)$.
%  \raft{Maybe clearer to first say $\E_D L(f_m(X), Y) = \inf_f L(f(X),Y)$ for all $m$, and then the easy implication of the convergence \jessie{Done}}
	However, we have $\E_X \propdis(\psi \circ f_m(X), D_X) = \propdis(\psi(u_m), p) = \propdis(\psi(u), p) \neq 0$, since $\psi(u) \not \in \gamma(p)$.
%  \raft{Almost -- just clarify that this is the expectation over $D$ too by construction \jessie{Kinda done?}}
	Thus, we observe $(L, \psi)$ is not consistent with respect to $\gamma$ (Definition~\ref{def:consistent-prop}).
\end{proof}

>>>>

We start by formalizing consistency in two ways that generalize across our four quadrants.
First, given a target loss $\ell$, we say $L$ is consistent if optimizing $L$ implies optimizing $\ell$ (Definition~\ref{def:consistent-ell}).
Second, given a target property $\gamma$, such as the $\alpha$-quantile, we say $L$ is consistent if optimizing $L$ implies approaching, in some sense, the correct statistic $\gamma(D_x)$ of the conditional distributions $D_x = \Pr[Y|X=x]$ (Definition~\ref{def:consistent-prop}).
We then observe that Definition~\ref{def:consistent-ell} is subsumed by Definition~\ref{def:consistent-prop}, and use this to show consistency implies $L$ indirectly elicits $\prop{\ell}$ or $\gamma$ respectively.

\begin{definition}[Consistent: loss]\label{def:consistent-ell}
	A loss and link $(L,\psi)$ are consistent with respect to a target loss $\ell$ if, for all distributions $D$ over input and label spaces $\X \times\Y$, and for all sequences of measurable hypothesis functions $\{f_m : \X \to \R\}$,
	\begin{align*}
	\E_D L(f_m(X), Y) \to \inf_f \E_D L(f(X), Y) &\implies \E_D \ell((\psi \circ f_m)(X), Y) \to \inf_f \E_D \ell((\psi \circ f)(X), Y)~.~
	\end{align*}
\end{definition}
%% Bo: resolved, Jun 1
%\botodo{On the right side, is measurable enough? I guess we need $\psi \circ f$ to be measurable...on the other hand, we don't want to constrain $f$ based on $\psi$. Do we need a condition on $\psi$ that guarantees $\psi \circ f$ is measurable?}
%\raft{I think okay as written -- if $\psi$ is restrictive, you won't achieve consistency.  And typically $f$ is much more expressive anyway, e.g. $\reals^\Y$ for classification.}

Instead of a target loss $\ell$, one may want to learn a target property, i.e. a conditional statistic such as the expected value, variance, or entropy.
In this case, following the tradition in the statistics literature on conditional estimation~\citep{gyorfi2006distribution,fan1998efficient,ruppert1997local},
we formalize consistency as converging to the correct conditional estimates of the property.
Convergence is measured by functions $\propdis(r, p)$ that formalize how close $r$ is to ``correct'' for conditional distribution $p$.
In particular we should have $\propdis(r,p) = 0 \iff r \in \gamma(p)$.
\btw{Bo: Would be nice to give some natural special cases: for a finite property, $\propdis(r,p) = \mathbf{1}[r \in \gamma(p)]$, and for single-valued properties with a distance metric on $\R$, $\propdis(r,p) = \text{dist}(r, \gamma(p))$.}

@@

%Lemma~\ref{lem:consistent-loss-implies-prop} confirms that consistency with respect to a target property is a strictly broader notion than with respect to a target loss.
Because each target loss elicits some property, but not all target properties can be elicited by a loss (e.g. the variance), consistency with respect to a property is the strictly broader notion.
This points to indirect elicitation as a natural necessary condition for consistency, as formalized in Theorem \ref{thm:consistent-implies-indir-elic}.

\begin{theorem}\label{thm:consistent-implies-indir-elic}
	For a surrogate $L \in \L$, if the pair $(L, \psi)$ is consistent with respect to a property $\gamma$ or a loss $\ell$ eliciting $\gamma$, then $(L, \psi)$ indirectly elicits $\gamma$.
\end{theorem}
\begin{proof}
%  \raft{Maybe ref the consistency definition here explicitly, at the top and the end}
	By Lemma~\ref{lem:consistent-loss-implies-prop}, it suffices to show the result for consistency with respect to a property $\gamma$, setting $\gamma := \prop{\ell}$ if $\ell$ is given instead.
	We show the contrapositive; suppose $(L, \psi)$ does not indirectly elicit $\gamma$, meaning we have some $p \in \simplex$ so that $u \in \Gamma(p)$ but $\psi(u) \not \in \gamma(p)$, where $\Gamma := \prop{L}$.
	Observe that we use the fact $\Gamma(p) \neq \emptyset$.
	Consider the constant sequence $\{f_m\}$ with $f_m(x) = u$ for all $m,x$, and take $D$ with full support on some $x \in \X$, and let $D_x = p$.
  Since $u \in \Gamma(p)$, we observe $\E_D L(f_m(X), Y) = \inf_f L(f(X),Y)$ for all $m$; in particular $\E_D L(f_m(X), Y) \to \inf_f L(f(X),Y)$.
%  \raft{Maybe clearer to first say $\E_D L(f_m(X), Y) = \inf_f L(f(X),Y)$ for all $m$, and then the easy implication of the convergence \jessie{Done}}
	However, we have $\E_X \propdis(\psi \circ f_m(X), D_X) = \propdis(\psi(u_m), p) = \propdis(\psi(u), p) \neq 0$, since $\psi(u) \not \in \gamma(p)$.
%  \raft{Almost -- just clarify that this is the expectation over $D$ too by construction \jessie{Kinda done?}}
	Thus, we observe $(L, \psi)$ is not consistent with respect to $\gamma$ (Definition~\ref{def:consistent-prop}).
\end{proof}
